---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# dojrace

<!-- badges: start -->

<!-- badges: end -->

## What is dojrace?

**dojrace** is a helper package for aggregating U.S. Census 2020 data into Department of Justice (DOJ) and Office of Management and Budget (OMB) racial categories using `tidycensus`.

### Why does this exist?

To conform to the DOJ & OMB definitions of racial categories. **If someone picks a racial category, they are counted in that group—even if they pick multiple racial categories.**\
- This means people may be double-counted.\
- “White” here means non-Hispanic white only.\
- This approach matches how the [DOJ and OMB](https://www.justice.gov/archives/opa/press-release/file/1429486/dl?inline=) define groups for enforcing the Voting Rights Act. \
- This allows voting strength and demographic analysis *without* excluding multiracial people, so that every person is reflected in the groups they select. \
- For more information, read [this story](https://jaxtrib.org/2021/12/16/jacksonvilles-redistricting-plans-ignore-federal-guidelines/) about a city that did not follow the OMB definitions.

## Installation

You can install the development version of dojrace from GitHub with:

```{r installation}
# install.packages("devtools")
devtools::install_github("apantazi/dojrace")
```

You’ll also need an API key from the U.S. Census Bureau and to set it with [tidycensus](https://github.com/walkerke/tidycensus):

```{r census api}
tidycensus::census_api_key("YOUR_API_KEY", install = TRUE)
```

## Example

This example computes DOJ/OMB racial categories for Duval County census blocks:

```{r example}
library(dojrace)
duval_blocks <- get_doj_race_cats(
  geography = "block",
  state = "FL",
  county = "Duval"
)
head(duval_blocks)
```

## What makes this different?

Most “race” aggregations from the Census are mutually exclusive and undercount multiracial residents.\
**dojrace** ensures every person is counted for every race they select, matching DOJ guidance, so you can analyze potential voting strength and compliance with the Voting Rights Act.

See [DOJ/OMB methodology](https://www.justice.gov/archives/opa/press-release/file/1429486/dl?inline=) for technical guidance.

```{r sessioninfo, echo=FALSE}
sessionInfo()
```

